Nitrogen cycling in the CLM5: controls on responses to environmental forcing

### 2018-10-04 main.tex

\documentclass[draft,linenumbers]{agujournal}
\draftfalse
\usepackage{hyperref}

\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
}

\journalname{Journal of Advances in Modeling Earth Systems (JAMES)}
\begin{document}

\authors{Rosie Fisher\affil{1},
Will Wieder\affil{1},
Ben Sanderson\affil{1},
Charlie Koven\affil{2},
Keith Oleson\affil{1},
Chonggang Xu\affil{3},
Ashehad Ali\affil{1},
Josh Fisher\affil{4},
Mingjie Shi\affil{4},
Katie Dagon\affil{1},
Danica Lombardozzi\affil{1},
Anthony Walker\affil{5},
S\"onke Zaehle\affil{6},
Gordon Bonan\affil{1},
David Lawrence\affil{1}
}

\title{Nitrogen cycling CLM5: controls on responses to environmental forcing}
\author{rosie fisher}
\date{April 2018}

\affiliation{1}{National Center for Atmospheric Research, Table Mesa Drive, Boulder, Colorado, USA}
\affiliation{2}{Lawrence Berkeley National Laboratory, Berkeley, California, USA}
\affiliation{3}{Los Alamos National Laboratory, Los Alamos, New Mexico, USA}
\affiliation{4}{NASA Jet Propulsion Laboratory, Pasadena, California, USA}
\affiliation{5}{Oak Ridge National Laboratory, Oak Ridge, Tennessee, USA}
\affiliation{6}{Max Planck Institute for Biogeochemistry, Jena, Germany}
\correspondingauthor{Rosie Fisher}{rfisher@ucar.edu}

\begin{keypoints}
\item A suite of modifications to the Nitrogen cycle representation were added to the Community Land Model (version 5).
\item New features of the N model include prognostic $V_{c,max}$ and $J_{max}$, representation of the carbon economic cost of Nitrogen uptake, switching of N uptake between alternative sources, variable tissue C:N ratios, dependant on N costs and distance from the target C:N ratio, and dynamic leaf resorption of Nitrogen. 
\item Here we assess the sensitivity of the model to a suite of parameters pertinent to the cycle of Carbon and Nitrogen. We assess the dependence of the model state to parameter uncertainty, and also the responses of the system to higher CO2, Nitrogen and temperature.
\item The model responses to CO2 and N fertilization are linked to the representation of the fraction of plants that can fix Nitrogen, and to the costs of N in the environment. Responses to temperature depend upon x,y,z. 
\item The model runs are conducted at representative sites in temperate, tropical and boreal systems.

@@

Simulating the cycling of nutrients, and how nutrient availability impacts ecosystem growth and function, has been repeatedly identified as a crucial element of Earth system models. In the CMIP5 analysis, only one model, the Community Land Model (v4.0) included an active nitrogen (N) cycle, and this model suggested that inclusion of N dynamics might significantly limit the ability of the terrestrial biosphere to respond to fertilization by increasing atmospheric carbon dioxide (\cite{friedlingstein2006},\cite{friedlingstein2014}, \cite{arora2013}, ).  This result raises the question of whether the inclusion of Nitrogen cycling necessarily generates such substantial limitations controls on the fertilization response of the biosphere.

Substantial uncertainty remains in the land surface modeling community regarding how nutrient limitations should be represented in models. Unlike representations of, say, photosynthetic responses to environmental conditions, or radiative transfer through forest canopies, our understanding of how nutrient cycling functions both within plants and in whole ecosystems is not governed by any well-tested and widely accepted paradigms.

Subsequent to CMIP5, many terrestrial biosphere and land surface models that represent N dynamics using alternative structural assumptions have been developed \cite{zaehle2010},  \cite{goll2012}, \cite{smith2014}, \cite{wang2007}). Detailed model-data synthesis activities, notably those conducted at North American free-air carbon dioxide enrichment (FACE) experiments, have revealed important differences in model behavior emerging in part from the theoretical framework for and parameterization of nutrient cycling processes (\cite{zaehle2014}).

The Community Land Model is the land surface representation within the Community Earth System Model (CESM, \cite{hurrell2013}.  Version 5 of the CLM was recently released to the research community as part of the CESM2.0 release. An overview of model developments is presented by \cite{lawrence2018}, a descriptions of the global Nitrogen cycling by Wieder et al. (in prep), the crop model (Lombardozzi et al. in prep), the land use components (Lawrence P et al. in prep), land surface hydrology (Swenson et al. in prep) and plant hydraulics (Kennedy et al. in prep).

The CLM5 model includes numerous major update of the nitrogen cycling representation subsequent to CLM4.0 and CLM4.5. The introduction of these new dynamic features into the CLM5 model has implications for parameter uncertainty and responses to environmental forcing. Wieder et al. (in prep) report the global-scale implications of the full model update.  In this paper, we describe the new components of the biogeochemical cycling of nitrogen, their impacts upon the model system state and responses to environmental forcing.

\subsection{Model Updates}
The CLM5 nitrogen cycling builds on the developments of the implementation of the CLM4.0 'CN' model (\cite{Thornton}, modifications to the biogeochemical cycling included in the interim CLM4.5 model (\cite{koven2013}, \cite{bonan2012}).  The new model contains updates to three different elements of the Nitrogen cycle, including: 
1)	The LUNA module, (Leaf Utilization of Nitrogen for Assimilation), described by \cite{xu2012} and \cite{ali2016}. 
2)	The FUN module (Fixation and Uptake of Nitrogen) described by \cite{fisher2010fun} \cite{brzostek2014} and \cite{shi2016}. 
3)	Part of the `FLEXCN' N cycle implementation, described by \cite{ghimire2016}.

\subsubsection{LUNA}
The LUNA model implements the prediction of $V_{c,max}$ and $J_{max}$ from environmental conditions.   The model determines the rate constants that are consistant with a co-limitation of photosynthesis by both electron capture and carboxylation processes. In this fashion, $V_{c,max}$ is largely independent of PFT, and is primarily controlled by the N per unit leaf area ($N_{area}$) and environmental conditions. The capacity of the LUNA model to represent appropriate responses to temperature and CO2 was assessed by \cite{xu2012} while the global calibration and the geographical predictions of the model were described by \cite{ali2016}.

\subsubsection{FUN}
The FUN model operates on the principle that for the primary sources of Nitrogen acquisition by plants (active uptake, retranslocation and symbiotic N fixation) there is a cost in terms of carbon. Further, the cost of each of these uptake pathways is variable in time and space, and thus plants will likely take up Nitrogen from pathways which are 'cheaper' to them. Where N is scarce in the environment, carbon that would otherwise have been used for growth can be deployed to acquire more N. The central element of FUN is a simulataneous equation which assumes, firstly, that the carbon available for growth ($C_{growth}$) is the total carbon available ($C_{avail}$), minus that used for Nitrogen uptake, $C_{nuptake}$:

\begin{equation}
C_{growth}=C_{avail}-C_{nuptake}
\end{equation}

Further, the Nitrogen acquired from the environment must equal that which is deployed in the growing plant tissues. The carbon used for growth is further reduced by the growth respiration term ($f_{gr}$) which applies only to tissues that are constructed:

\begin{equation}
C_{nuptake}/N_{cost} =\frac{C_{growth}}{(CN_{target}/(1.0+f_{gr})}
\end{equation}

Thus the Carbon that is expended on N uptake is determined by:
\begin{equation}
C_{nuptake} =\frac{C_{avail}}{ ( (1.0+f_{gr})*(CN_{target} / N_{cost}) + 1) }
\end{equation}

@@

\subsubsection{FLEXCN}
\cite{ghimire2016} implemented a suite of modifications to the Nitrogen cycling including variable tissue C:N ratio, prognostic Vcmax (resulting from this dynamic leaf N content) and also an alternative Nitrogen uptake algorithm. In CLM5, the LUNA model predicts $V_{c,max}$ from leaf N content (while also taking into account environmental conditions), and the FUN model represents Nitrogen uptake, the primary element of CLM5 utilized from the \cite{ghimire2016} model is the modification to the allocation scheme that allows tissue C:N ratios to be modified in response to varying supply of Nitrogen. In this new algorithm, Nitrogen supply is partitioned between tissues according to their relative 'demand' terms (ascertained from the target C:N ratio and the carbon allocated to the pool) and C:N ratios vary with Nitrogen supply rates. This new allocation code we term the FLEXCN module, and it is documented in ***.

\subsubsection{Adaptation of FUN to flexible CN ratios}
The FUN model was originally conceived with a static tissue CN ratio, and will by default allocated an amount of carbon to N uptake that will exactly track the target CN ratio. In CLM5, one of the aims of the development process was to incorporate a varying CN ratio. To that end, it is necessary to modify the amount of Carbon expended on N uptake (and thus the amount of N required). The degree to which tissue CN content varies as N becomes limiting in unclear in the literature, and N cycle models typically include heuristic representations of tissue N adjustments in light of limitations (including \cite{zaehle2010},\cite{ghimire2016}). Here we include a placeholder algorithm that adapts to 1) the environmental cost of N acquisition and 2) the degree to which the plant is already far from it's target N content. When N is expensive in the environment, less C is spent on uptake, increasing CN ratios where N is scarce. When plants become extremely N limited, however, expenditure increases again to prevent CN ratios becoming unrealistically high.

\subsubsection{Justification for model developments}
Several elements of model behaviour that are critical for simulation of plant responses to variation in nitrogen availability  emerged from the FACE model data synthesis activity that were not captured by previous versions of the CLM Nitrogen cycle (\cite{zaehle2014}, \cite{medlyn2015using}). For example, changes in the C:N ratio accounted for large changes in ecosystem carbon storage,  but these are not captured by the static tissue C:N ratios used in the CLM4.0 and CLM4.5 nitrogen cycle. For this reason, we introduce a flexible tissue C:N ratio into the CLM5.

Most land surface models predict photosynthetic parameters from leaf nitrogen content (\cite{kattge2009}, \cite{bonan2012}). There is significant evidence, however, that plant photosynthetic capacity can respond to environmental conditions, including CO$_{2}$ fertilization \cite{ainsworth2007} and changing temperature \cite{hikosaka2005}, irradiance (\cite{niinemets1998}) and soil moisture (\cite{keenan2009}). Given the importance of photosynthetic capacity as a key model parameter (\cite{rogers2017}) the use of a static photosynthetic capacity appears to be a poor means of capturing its variation in time and space and responses to climatic change (\cite{walker2017}).

\subsubsection{Content of this paper}

@@

Four 
We ran the default model configuration to equilibrium, and

@@

\section{Discussion}

\subsubsection{Discourse on how target CN ratio could be optimized but isn't}
\cite{thomas2014}, \cite{friedlingstein1999}, \cite{franklin2012}, \cite{mcmurtrie2013}, \cite{anten2011}, \cite{vanwijk2003}

\subsubsection{discuss geographical spread of predictions for N and CO2 fertilizaion}
\cite{hickler2008}

>>>>

\documentclass[draft,linenumbers]{agujournal}

\draftfalse

\usepackage{hyperref}
\hypersetup{
colorlinks=true,
linkcolor=blue,
filecolor=magenta,
urlcolor=cyan}

\journalname{Journal of Advances in Modeling Earth Systems (JAMES)}

\begin{document}

\authors{Rosie Fisher\affil{1},
Will Wieder\affil{1},
Ben Sanderson\affil{1},
Charlie Koven\affil{2},
Keith Oleson\affil{1},
Chonggang Xu\affil{3},
Ashehad Ali\affil{1},
Josh Fisher\affil{4},
Mingjie Shi\affil{4},
Katie Dagon\affil{1},
Danica Lombardozzi\affil{1},
Anthony Walker\affil{5},
Gordon Bonan\affil{1},
David Lawrence\affil{1}
}

\title{Nitrogen cycling in the CLM5: controls on responses to environmental forcing}
\author{rosie fisher}
\date{April 2018}

\affiliation{1}{National Center for Atmospheric Research, Table Mesa Drive, Boulder, Colorado, USA}

\affiliation{2}{Lawrence Berkeley National Laboratory, Berkeley, California, USA}

\affiliation{3}{Los Alamos National Laboratory, Los Alamos, New Mexico, USA}

\affiliation{4}{NASA Jet Propulsion Laboratory, Pasadena, California, USA}

\affiliation{5}{Oak Ridge National Laboratory, Oak Ridge, Tennessee, USA}

\correspondingauthor{Rosie Fisher}{rfisher@ucar.edu}

\begin{keypoints}

\item A suite of modifications to the Nitrogen cycle representation were added to the Community Land Model, version 5 (CLM5).

\item New features of the N model include prognostic $V_{c,max}$ and $J_{max}$, representation of the carbon economic cost of Nitrogen uptake, switching of N uptake between alternative sources, variable tissue C:N ratios and dynamic leaf resorption of Nitrogen.

\item Here we assess the sensitivity of the model to a suite of parameters pertinent to the cycle of Carbon and Nitrogen. We assess the dependence of the both the model state, and also the responses of the system to higher CO2, Nitrogen and temperature to parameter uncertainty. Model runs are conducted at representative individual sites in temperate, tropical and boreal systems.

\item The model responses to CO2 and N fertilization are primarily linked to the representation of the fraction of plants that can fix Nitrogen, and to the costs of N in the environment.

\item Identification of the primary factors driving ecosystem responses to fertilization both assists in the interpretation of global simulations and in the identification of priorities for fieldwork and process updates.

@@

Simulating the cycling of nutrients, and how nutrient availability impacts ecosystem growth and function, has been repeatedly identified as a crucial element of Earth system models (\cite{piao2013}, \cite{gruber2008}, \cite{wang2009}). In the CMIP5 multi-model intercomparison process, only one model, the Community Land Model v4.0, included an active nitrogen (N) cycle. The CLM4.0 projected that inclusion of N dynamics might significantly limit the ability of the terrestrial biosphere to respond to fertilization by increasing atmospheric carbon dioxide (\cite{friedlingstein2006},\cite{friedlingstein2014}, \cite{arora2013}). Subsequent and parallel to CMIP5, many terrestrial biosphere and land surface models that represent N dynamics using alternative structural assumptions have been developed \cite{wang2007}, \cite{zaehle2010}, \cite{goll2012}, \cite{smith2014}). Nonetheless, substantial uncertainty remains in the land surface modeling community regarding how nutrient limitations should be represented in models (\cite{zaehledalmonech2011}).

Unlike representations of photosynthetic responses to environmental conditions, or radiative transfer through forest canopies, our understanding of how nutrient cycling functions both within plants and in whole ecosystems is not governed by any well-tested and widely accepted paradigms. Detailed model-data synthesis activities, notably those conducted at North American free-air carbon dioxide enrichment (FACE) experiments, have revealed important differences in model behavior emerging in part from the absence of a broadly accepted theoretical framework for construction and parameterization of nutrient cycling processes (\cite{zaehle2014}). Few, if any studies, however, have looked at how CO$_{2}$ or N fertilization are governed by the parameters or structure of a given model.

The Community Land Model is the land surface representation within the Community Earth System Model (CESM, \cite{hurrell2013}). Version 5 of the CLM was recently released to the research community as part of the CESM2.0 release. An overview of model developments is presented by \cite{lawrence2018}, a descriptions of the global Nitrogen cycling by Wieder et al. (in prep), the crop model (Lombardozzi et al. in prep), the land use components (Lawrence P et al. in prep), land surface hydrology (Swenson et al. in prep) and plant hydraulics (Kennedy et al. in prep).

The CLM5 model includes numerous major update of the nitrogen cycling representation subsequent to CLM4.0 and CLM4.5. The introduction of several new dynamical features into the CLM5 has implications for both parameteric uncertainty and responses to environmental forcing. Wieder et al. (in prep) report the global-scale implications of the full model update and note that the CLM( responds more strongly to CO2 and less strongly to N fertilization than its predecessors. In this paper, we describe the new components of the biogeochemical cycling of nitrogen, their impacts upon the model system state and responses to environmental forcing.

\subsection{Model Updates}

The CLM5 nitrogen cycling builds on the developments of the implementation of the CLM4.0 'CN' model (\cite{thornton2007}), and on modifications to the biogeochemical cycling (notably vertically stratified soil decomposition processes) included in the interim CLM4.5 model (\cite{koven2013}, \cite{bonan2012}). The new model integrates three major additional prognostic elements of the Nitrogen cycle, including: \\

1) The `LUNA' model (Leaf Utilization of Nitrogen for Assimilation), which simulates distribution of N between different leaf assimilation processes (\cite{xu2012}, \cite{ali2016}) \\

2) The `FUN' module (Fixation and Uptake of Nitrogen), which simulates the dynamics of Nitrogen acquisition from the environment (\cite{fisher2010fun}, \cite{brzostek2014}, \cite{shi2016}).\\

3) The `FLEXCN' N cycle implementation, which is used here primarily to allow variation in tissue C:N ratios, and is adapted from \cite{ghimire2016}.\\

More details on each of these new components is given below. A full technical description of the CLM5 is an appendix to \cite{lawrence2018} and is also available online at \url{(https://escomp.github.io/ctsm-docs/doc/build/html/tech_note/index.html).}

\subsubsection{LUNA}

Most land surface models predict photosynthetic parameters from leaf nitrogen content (\cite{kattge2009}, \cite{bonan2012}). There is significant evidence, however, that plant photosynthetic capacity can respond to environmental conditions, including CO$_{2}$ fertilization (\cite{ainsworth2007}) and changing temperature (\cite{hikosaka2005}), irradiance (\cite{niinemets1998}) and soil moisture (\cite{keenan2009}). Given the importance of photosynthetic capacity as a key model parameter (\cite{rogers2017}) the use of a static photosynthetic capacity appears to be a poor means of capturing its variation in time and space and responses to climatic change (\cite{walker2017}).

The LUNA model predicts the optimal balance of of $V_{c,max}$ and $J_{max}$ for the prevailing time-averaged environmental conditions. The model determines the rate constants that are consistent with a co-limitation of photosynthesis by both electron capture and carboxylation processes. In this fashion, $V_{c,max}$ is primarily controlled by the N per unit leaf area ($N_{area}$, which is itself a function of the target C/N ratio, the specific leaf area, and the prognostic variation in leaf C/N ratios) and environmental conditions. The capacity of the LUNA model to represent appropriate responses to temperature and CO$_{2}$ was assessed by \cite{xu2012} while the global calibration and the geographical predictions of the model were described by \cite{ali2016}.

\subsubsection{FUN}

Under circumstances where insufficient N exists to match all of the carbon assimilated for a given C:N ratio, the CLM4.0 model reduced the gross photosynthetic flux down to the level at which growth could be supported by the assimilated N. This process occurred after the calculation of stomatal conductance (which is linked to assimilation rates via the Ball-Berry model) and therefore created inconsistencies between C assimilation and water cycling, especially under conditions of N limitation, as discussed extensively in the literature (\cite{medlyn2011}, \cite{bonan2012}, \cite{dekauwe2014}, \cite{walker2014}).

One major issue confronting N cycle models is how to deal with this 'excess' carbon under N limiting conditions (\cite{zaehle2010}, \cite{dekauwe2014}). The FUN model operates on the principle that for the alternative sources of Nitrogen acquisition by plants (active uptake from soil, retranslocation from senescent tissues, and symbiotic N fixation) there is a concurrent cost in terms of carbon. Further, the cost of each of these uptake pathways is variable in both time and space. FUN hypothesizes that plants will take up Nitrogen from pathways which are 'cheaper' to them. Where N is scarce in the environment, carbon that would otherwise have been used for plant growth can be deployed to acquire (pay for) more N. The central element of FUN is a simultaneous equation which assumes, firstly, that ($C_{avail}$), the total carbon available (after maintenance respiration and replenishment of stores are accounted for) is either used for plant growth ($C_{growth}$) or for Nitrogen uptake, ($C_{nuptake}$).

\begin{equation}
C_{growth}=C_{avail}+C_{nuptake}
\end{equation}

Further, the Nitrogen acquired from the environment ($N_{uptake}$) must equal that which is deployed in the growing plant tissues. The carbon used for growth is further reduced by the growth respiration term ($f_{gr}$) which applies only to tissues that are constructed:

\begin{equation}
N_{uptake}=C_{nuptake}/N_{cost} =\frac{C_{growth}}{(CN_{target}/(1.0+f_{gr})}
\end{equation}

Thus the Carbon that is expended on N uptake is determined by:

\begin{equation}
C_{nuptake} =\frac{C_{avail}}{ ( (1.0+f_{gr})*(CN_{target} / N_{cost}) + 1) }
\end{equation}

@@

A further issue addressed by FUN is the preferential use of symbiotic N fixation when soil mineral N concentrations are low (\cite{vitousek2002}). The CLM4.0 and CLM4.5 predict N fixation as a linear function of net primary productivity (\cite{wieder2015} illustrate the impacts of uncertainty in this function) but FUN considers fixation rates to be an emergent property of plant assimilation rates (providing carbon for fixation), the relative costs of environmental N acquisition, and temperature (which exerts a primary control over the enzymatic processes, following \cite{houlton2008}). Fixation is preferred when the costs of N acquisition are higher than that of fixation. Note that while the CLM5 has no representation of biological succession (\cite{fisher2018}), this property nonetheless affects the amount of time taken to achieve a spun-up state from bare ground, in productive areas, N fixation is preferred and soil N builds up much quicker. In low productivity areas, carbon is required to fix N from a 'col start' initial state, impacting the survival of marginal plants; thus higher initial stocks of N are required to allow growth once metabolic and turnover costs have been met in early spin-up.

\subsubsection{FLEXCN}

Several elements of model behaviour that are critical for simulation of plant responses to variation in nitrogen availability emerged from the FACE model data synthesis activity that were not captured by previous versions of the CLM Nitrogen cycle (\cite{zaehle2014}, \cite{medlyn2015using}). For example, changes in the C:N ratio accounted for large changes in ecosystem carbon storage, but these are not captured by the static tissue C:N ratios used in the CLM4.0 and CLM4.5 nitrogen cycle. For this reason, we introduce a flexible tissue C:N ratio into the CLM5.

\cite{ghimire2016} implemented a suite of modifications to the CLM Nitrogen cycling model, including variable tissue C:N ratio, prognostic $V_{c,max}$ (resulting from this dynamic leaf N content) and also an alternative Nitrogen uptake algorithm. In the default configuration of CLM5, the LUNA model predicts $V_{c,max}$ from leaf N content (while also taking into account environmental conditions), and the FUN model represents Nitrogen uptake, the primary element of CLM5 utilized from the \cite{ghimire2016} model is the modification to the allocation scheme that allows tissue C:N ratios to be modified in response to varying supply of Nitrogen (although the alternative schemes implemented by \cite{ghimire2016} are available as options). In this new allocation algorithm, the total Nitrogen supply in each timestep is partitioned between tissues according to their relative 'demand' terms (ascertained from the target C:N ratio and the carbon allocated to the pool) and C:N ratios vary with Nitrogen supply rates. This new allocation code we term the FLEXCN module.

\subsubsection{Adaptation of FUN to flexible CN ratios}

The FUN model was originally conceived with a static tissue CN ratio, and will by default allocated an amount of carbon to N uptake that will exactly track the target CN ratio. In CLM5, one of the aims of the development process was to incorporate a varying CN ratio. To that end, it is necessary to modify the amount of Carbon expended on N uptake (and thus the amount of N required). The degree to which tissue CN content varies as N becomes limiting in unclear in the literature, and N cycle models typically include heuristic representations of tissue N adjustments in light of limitations (including \cite{zaehle2010},\cite{ghimire2016}). Here we include a placeholder algorithm that adapts to 1) the environmental cost of N acquisition and 2) the degree to which the plant is already far from it's target N content. When N is expensive in the environment, less C is spent on uptake, increasing CN ratios where N is scarce. When plants become extremely N limited, however, expenditure increases again to prevent CN ratios becoming unrealistically high. This new algorithm is also documented in the CLM5 technical description ((\url{(https://escomp.github.io/ctsm-docs/doc/build/html/tech_note/index.html)}) and includes three tuning parameters ($flexcn_{a}$, $flexcn_{b}$ \& $flexcn_{c}$) the sensitivity of which is investigated here.

\subsection{Content of this paper}

In this paper, we assess in detail the sensitivity of the CLM5 model structure to its input parameters. We use\\

1. A range of single site experiments \\

2. One-at-a-time and global parameter uncertainty analyses\\

3. CO$_{2}$ fertilization experiments and\\

4. Nitrogen addition experiments.\\

to fully investigate the performance of the model under steady-state and step-change modifications in forcing. Given the large amount of analysis and attention devoted to assessing the impacts of transient CO$_{2}$ and N in the context of global feedback loop analysis, we thus consider it a worthwhile exercise to illustrate how parametric choices in the CLM5 might affect overall behaviour in larger-scale experiments, including the controls on global carbon fertilization `$\beta$' values. We also hope to highlight areas of process and parameter uncertainty that appear to require greater attention in future model development and testing cycles.

Because of the increased number of prognostic processes with the CLM5, many alternative representation of C:N cycle interactions are possible within the same model structural space, including an effective removal of N fixation capacity (where carbon available for fixation is fixed at 0$\%$ of assimilation) and variation in the degree to which to C:N ratios are flexible. We further investigate the consequences of turning off the LUNA model and returning to the fixed leaf N/$V_{c,max}$ relationship from \cite{ghimire2016}.

@@

\subsection{Sites}

For this experiment we utilized six different experimental sites. Two in tropical forests (Caxiuana and Barro Colorado Island) two in temperate forest (the Oak Ridge National Lab - Free Air Carbon Dioxide Enrichment (FACE) experiment and Harvard Forest), one in high altitude forest (Niwot Ridge, Colorado) and one in seasonally dry evergreen forest (Metolius Forest).

\subsection{Simulation Protocol}

We spun up the default version of the model for 400 years at each site, in 'accelerated spin-up' mode (where the turnover of slower cycling carbon and nitrogen pools is accelerated, see (/cite{lawrence}) using pre-industrial CO$_{2}$ concentrations (274ppm) and recycling the available site-level meteorological inputs. 400 years was sufficient to bring all carbon and nitrogen pools into equilibrium. Subsequent to this, parameters were perturbed (see below), and a second spin-up conducted for 180 years at each site, with accelerated mode turned off. Using the state at the end of the perturbed parameter spin-ups, we ran transient simulations for each site and parameter combination starting in 1850 and invoking transient CO$_{2}$ and Nitrogen deposition until the end of the observed meteorological forcing. We ran additional CO$_{2}$ and Nitrogen deposition experiments by increasing the CO$_{2}$ concentration to 550ppm. For the Oak Ridge FACE site, we only increased CO$_{2}$ in the growing season, in line with the experimental protocol applied there and using the observed CO$_{2}$ time-series. For all other sites, we increased CO$_{2}$ to 550ppm year round starting in 1997**. For Nitrogen, we added 5kgN $m^{-2}$/year, also starting in 1997, so that the results can be compared with those reported by Wieder et al. (in prep).

\subsection{Perturbed Physics Ensemble (PPE)}

The CLM5 has a very large number of parameters to which the C and N cycles are potentially sensitive. Here we focused on parameters of particular relevance to the C and N cycles and on those which are newly introduced by the modified N cycle. We choose a larger set of parameters for one-at-a-time analyses, followed by a smaller set of these to which we subject the model to global sensitivity tests. Knowledge of model structure as well as previous investigations of the CLM model informs our choice of parameters. Our principle goal is to identify parameters that might have large impacts on responses to model forcing, as opposed to changes in the base state of the model, since these are the primary determinants of the carbon cycle feedback in the context of the Earth System Model. However, we also report responses of the model state to perturbations. Additional investigations into the parametric sensitivity of the hydrology and surface energy balance in the CLM5 are planned (Dagon et al. in prep) and into the plant hydraulics parameterizations (Kennedy et al in prep), thus, we have not included the full range of those parameters here.

In the first instance, we conduct one-at-a-time (OAAT) perturbations of the parameters around a mean state, to allow a straightforward visualization and interpretation of the results. For each parameter, for each site, we conduct four simulations, with two above and two below the default value (as well as the default simulation).

\subsubsection{Parameter Choices}

The parameters included in our sensitivity analysis (Table \ref{table_parameters}) are: \\

1) \textbf{Specific Leaf Area}, \emph{`SLA'}, which is a critical parameter determining the area of canopy (in m$^{2}$) derived from one gram of leaf biomass. This parameter is known to be one of the primary elements of the leaf economics spectrum (\cite{wright2004}), and is well-characterized by the TRY database (\cite{kattge2011}).\\

2) \textbf{Fine root area per unit leaf area}, \emph{`froot\_leaf'}, the ratio of fine root biomass to leaf area. Unlike in previous versions of the CLM, fine root biomass now affects the capacity of the model to acquire both water and nutrients (via the resistance to water flow and the cost of N acquisition), thus the impact of this parameter is of greater interest than it's previous functionality as a simple 'tax' on plant growth. \\

3) \textbf{Stem to leaf area ratio}, \emph{`stem\_leaf'}, reflects the balance between accumulation of stem and woody biomass, and leaf area. Increases in this parameter decrease the leaf area index achievable for a given productivity, and conversely increase the equilibrium woody biomass predictions.\\

4) \textbf{Nitrogen Costs}, \emph{`n\_costs'}, are a set of parameters that determine the environmental cost of Nitrogen uptake from the soils. These six parameters were previously defined by \cite{brzostek2014}. Here and in CLM5 in general, we maintain the ratios of the parameters as defined by the existing FUN model calibration, but then allow the magnitudes of the parameters to change concomitantly, since the definitions of soil N availability vary between CLM5 and \cite{brzostek2014}. These parameters include \emph{ekc\_active} and \emph{ekn\_active}, the rate constants that relate the cost of ectomycorhhizal uptake to root carbon (ekc) and soil N concentration (ekn), and their counterparts \emph{akc\_active} and \emph{akn\_active} for arbuscular mycorrhizal fungi, and \emph{kc\_nonmyc} and \emph{kn\_nonmyc} for non-mycorrhizal active uptake. \\

5) \textbf{The fraction of carbon that can be used for fixation}, \emph{`'fracfixers'}. This parameter is the fraction of the assimilation (gross primary productivity - autotropihic respiration, or $gpp$ - $R_{auto}$) that can be used for symbiotic N fixation by the FUN model, and is a proxy for the ecosystem-level fractional productivity of Nitrogen fixing plants, in lieu of, for example, a prognostic model of N fixer distributions, or an input map of their abundance, both of which could plausibly be added in future generations of CLM. \\

6) \textbf{The fraction of carbon spent on respiration per unit of new tissue (growth respiration)}, \emph{grperc}, was revised substantially in CLM5 based on Atkin O. (pers comm.) from 0.3 to 0.11. Here we test the impacts of this change on overall carbon budgets.\\

7) \textbf{Leaf carbon:nitrogen target}, \emph{leafcn}. CLM5 uses the concept of a 'target' leafcn ratio, around which flexibility is allowed, given the environmental costs of Nitrogen. LeafCN ratios are used also as input to the LUNA model to determine photosynthetic rate constants ($V_{c,max}$ and $J_{max}$) \\

8) \textbf{Slope of stomatal conductance model}, \emph{medlyn\_slope}. This new parameter, which replaces the fixed Ball-Berry slope used in CLM5, (and is PFT-dependant), is critically linked to the response of ecosystems to CO$_{2}$, assessed in this set of experiments. \\

9) \textbf{Respiration intercept}, \emph{lmr\_intercept}. The new leaf maintenance respiration model, proposed by \cite{atkin2015} and adopted by CLM5, has it's intercept as a PFT dependant parameter, with variation between PFTs reported by \cite{atkin2015} and used herein. Here we test the impacts of the uncertainty in the observations of this parameter. \\

10) \textbf{Fraction of uptake that is ectomycorrhizal}, \emph{frac\_ecto}, is another ecosystem-level property defined by the FUN model and allowing the proportions of the fungal symbiotic pathways to be modified. Default parameters are as provided by \cite{shi2016}. This capacity was not fully explored in the CLM5 development process and so here we assess its importance. \\

11,12,13) \textbf{Parameters a,b and c of the flexible CN model linked to FUN}, \emph{cn\_flex\_a, cn\_flex\_b, \& cn\_flex\_c}, as described in Appendix 1, which determine the response of tissue CN ratios to depletion of N in both the environment and the plant tissues themselves.

\subsubsection{Parameters not used}

We conducted a set of tests on the parameters of the denitrification model, to assess the impacts of differing rates of N loss on the system. However, recent analysis has shown that the rates of overall denitrification in CLM5 are very low, overall (compared to plant uptake and immobilization \textbf{I've forgotten why this is? Can someone remind me}) and hence these parameters had very little impact. We did not conduct tests on leaf lifespan, since for evergreen trees in CLM5, leaf lifespan and specific leaf area have very similar impacts on overall achievable leaf area, and for deciduous trees, leaf lifespan is primarily determined by phenological triggers. We did not investigate any parameters relating to the crop model (Lombardozzi et al. in prep) nor to plant harvest, fire, dust, soil biogeochemical cycling (the latter being is largely unaltered from CLM4.5, \cite{koven2013} apart from modifications presented by \cite{lawrence}.

\subsubsection{Parameter Ranges}

\emph{Specific Leaf Area} and \emph{leafcn} are defined in the TRY database, and their distributions are log-normally distributed. Thus, we take the average log-normal standard deviation range from TRY and apply it across all PFTs. Thus the lower bounds are closer to the mean than the higher bounds (Table \ref{table_ranges}). The range of \emph{fracfixers} and \emph{frac\_ecto} were set to vary across the whole logical range from 0 to 1 (0 to 100\% of ecosystem productivity being available for fixation or ectomycorrhizal uptake). \emph{grperc} ranges from its logical lower bound of zero up to the range of the previously used CLM value (0.3). \emph{lmr\_intercept} and \emph{medlyn\_slope} are both set such that the range of values tested represents the range across all the PFTs reported in \cite{atkin2015} and \cite{dekauwe2015}. The ranges of the \emph{cn\_flex} parameters are unknown, since these physiological controls are poorly understood, thus we used ranges that span the response limits of these parameters, particularly for \emph{cn\_flex\_c} which we explored across four orders of magnitude to investigate the impacts of leafCN flexibility. The allometric parameters ***. Nitrogen parameters.

\subsection{Analysis}\\
We calculated the impact of the CO$_{2}$ and N deposition by assessing the ratio between the control and the fertilized simulations for the years 2001-2003. For some parameter settings, the constraint on growth placed the system in a state close to death. Addition of N or CO$_{2}$ therefore caused either a very strong recovery of the plants, or no recovery at all where biomass is already zero, leading to occasional large non-linearities in how parameters are linked with responses to environmental forcing. Therefore, for those parameter combinations which cause LAI to be less than 80\% of the control simulation value, we exclude the CO$_{2}$ and Nitrogen responses from our plots, so as to focus on the responses of parameter combinations that give realistic initial vegetation conditions.

@@

The analysis conducted involved 6 sites, with 13 parameters perturbed to 4 different values, the each of those with an increase CO2, N and control run. Including the default simulations, this is 936 different run configurations. In this analysis, we look at the parameteric responses of five chosen output variables (GPP, NPP, LAI, Leaf N and vegetation carbon) first for their impacts on the system state and then on the responses to carbon and nitrogen limitation.

\subsection{Parametric control over system state}
Figure *** shows the one-at-a-time parameter responses of GPP, NPP, LAI, Leaf N content and the carbon spent on N uptake at the Caxiuana site. Analagous figures for the other five sites are in the SI figures 1-5 ***.

One notable feature of the sensitivity analysis here is the difference in the dominant parameter responses between different output variables, particularly as one moves along the C allocation cascade from GPP, to NPP and then LAI. Leaf CN ratio, (\emph{leafcn}) for example, is a dominant control over GPP, but is much less important for NPP, where the respiration intercept ($lmr\_{intercept}$), fine root ratio, and the fraction of fixers (by means of the carbon expenditure on respiration) impart a greater impact. For leaf area index (LAI), the allocation parameters, specific leaf area, root:leaf and stem:leaf ratio, have a much greater import, than almost all the parameters that affect NPP, given the linear and fixed allocation scheme. We here discuss the causality of how each individual parameter affects the major output variables, and how these responses relate to model structure, particularly in cases where the response outputs are not intuitive.
